### 2014-10-27 opdracht_1/beamer.tex
vergeten te committen

% Created 2014-10-27 Mon 11:11
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{hyperref}
\tolerance=1000
\graphicspath{{graphics/}}
\usepackage[style=authoryear,hyperref,backref,square,natbib,ibidtracker=false]{biblatex}
\bibliography{bibliography}
\usepackage[english]{babel}
\usepackage{graphicx}
\usetheme[height=20pt]{kuleuven}
\author{Ward Schodts, Xavier Goás Aguililla}
\date{maandag 10 november 2014}
\title{Internet of Things code deployment metrics}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 24.3.1 (Org mode N/A)}}
\begin{document}

\maketitle
\begin{frame}{Outline}
\tableofcontents
\end{frame}

\selectlanguage{english}

\section{Introduction}
\label{sec-1}
\begin{frame}[label=sec-1-1]{wireless sensor networks}
\end{frame}
\begin{frame}[label=sec-1-2]{sample applications}
\end{frame}
\begin{frame}[label=sec-1-3]{store, compute, transmit?}
\end{frame}
\section{Middleware for WSNs}
\label{sec-2}
\begin{frame}[label=sec-2-1]{what is middleware?}
\end{frame}
\begin{frame}[label=sec-2-2]{approaches to middleware}
\end{frame}
\begin{frame}[label=sec-2-3]{LooCi}
\end{frame}
\section{Evaluating energy use}
\label{sec-3}
\begin{frame}[label=sec-3-1]{oscilloscopy}
\end{frame}
\begin{frame}[label=sec-3-2]{accuracy}
\end{frame}
\section{Conclusion}
\label{sec-4}
% Emacs 24.3.1 (Org mode N/A)
\end{document}