### 2015-03-24 tussentijdsverslag/tussentijdsverslag.tex
tussentijdsverslag

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Tussentijds verslag WV}
\author{Ward  Xavier}
\date{March 2015}

\begin{document}
\maketitle

\section{Update}

\section{Voorlopige resultaten}

\section{Vragen}

\end{document}