### 2015-03-24 tussentijdsverslag/tussentijdsverslag.tex
interim report

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\title{Interim Report WV}
\author{Ward  Xavier}
\date{March 2015}

\begin{document}
\maketitle

\section{Update}
Due neverending problems with over-the-air-deployment we decided to stop researching using Looci. We are now using just Contiki on the motes. We deploy components with the flasher.
\section{Interim Results}
\subsection{RAM}

\subsection{Flash}
As discussed with Klaas, we will only research this option when we have time.

\subsection{Interim Conclusion}
\section{Questions}

\end{document}